Missing links and the topological robustness of food webs

### 2022-09-20 manuscript/ms_C3_v2.Rmd
Manuscript for feedback to Owen

---
title: Missing links, forbidden links, and the topological robustness of food webs
authors:
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: anubhav.gupta@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Anubhav Gupta
  thanks: Corresponding author
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: owen.petchey@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Owen L. Petchey
geometry: margin=1in
abstract: |

  1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food web. Undersampling can be compensated for by using food web models such as allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occuring due to primary extinctions) increases with connectance (the number of trophic links divided by the number of possible links), and (ii) that predicted food webs usually have greater connectance than observed ones. Thus we expect that predicted food webs are more robust than observed ones. This expectation has never, to our knowledge, been tested, nor has the effect size been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links, though can also make false positives. We did this for 12 different food webs from a wide variety of ecosystems.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, and this can be attributed to the higher connectance of the predicted food webs. On average, for every X unit of increase in connectance, we found the food webs to be robust by YY units for random extinction scenario.**OP: Here something about the effect size.**
  5)  These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, and so it would be interesting to compare the dynamical stability of observed and predicted food webs, as well as the topological stability that we have focused on.

@@

## Background on anthropogenic changes and its impact on food web

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity, and can lead to food web collapse [@ullah2018]. This collapse in the food web is due to cascades of secondary extinctions in a food web because of primary loss of species, for example due to habitat destruction. The rate of collapse of these predicted food webs is dependent on the structure and complexity of those food webs [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. Therefore, research focused on cascading secondary extinctions also known as 'community viability analysis' have been performed extensively in the past few decades [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005].

## Briefly explain work done by Jennifer Dunne on food web [OP] robustness to primary extinctions, and also what has been done since, particularly concerning the importance of connectance

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. These studies showed that the robustness of the food webs increases with food web connectance. Also, the removal of the most connected species cause considerably more secondary extinctions than random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. These studies provide an alternate solution to investigate the impact of primary extinctions in a food web when canonical experiments in natural ecosystems are not possible.

Along with robustness based on topological structure of a food web, robustness based on the food web dynamics has been studied as well (reference). Topological approaches only require the food web structure whereas dynamical approaches also require the temporal dynamics of the food web along with the food web structure. For example: @williamsEffectsNetworkDynamical2008 combined models of network structure with models of bionergetic dynamics to study the role of food web topology and nonlinear dynamics on species coexistence in complex ecological networks.

[OP] Explain (either here or in the methods) difference between topological and abundance (given by dynamics) based criteria for a secondary extinction occurring.

## [OP] State what is the problem...

A key assumption of the observed food webs is that the food webs are very well sampled i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and then do not represent all of the feeding links that occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort as compared to others whereas some trophic links remain unobserved because of linkage constraints irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. One solution to this problem is to use a food web model such as Allometric Diet Breadth Model (ADBM) to predict which are the missing links, and to then measure the robustness of the predicted food web. ADBM is a mechanistic model constructed using rules based on body sizes of prey and predator where trophic interactions satisfying those rules would be predicted by the model which at the same time are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links, and also may predict some links that could not, in fact occur.

## What we do in this study

In our study, we investigate the robustness of the ADBM predicted food webs as compared to the observed food webs and quantify the effect of overestimation of connectance on the robustness of these predicted food webs. We do this by simulating primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

It is crucial to investigate the implication of this consistent overestimation of connectance in the robustness of predicted food webs. We expect that the ADBM predicted food webs would be more robust as compared to the observed food webs, and for the greater robustness to be related to the amount by which the ADBM overestimates connectance. In this study, we simulate primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

[OP] ## Provide overview of the methods

In the upcoming sections, we present a detailed account of the implementation of simulation of primary extinctions for three different scenarios on 12 food webs predicted by the ADBM from wide variety of ecosystems, and compute the resultant secondary extinctions. We then compute a robustness metric to quantify the robustness of those predicted food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables in the model are: energy content of prey, handling times of the predator on prey, space clearance rate of predator on prey, and prey densities. All are derived from the body sizes of the species via allometries.

@@

The observed food webs that we fit the ADBM to belong to marine, freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). The observed connectance of these food webs is from 0.03 to 0.24 and there are 29 to 239 species. The food webs contain primary producers, herbivores, carnivores, parasites, and parasitoids. They also contain various types of feeding interactions, including predation, herbivory, bacterivory, parasitism, pathogenic, and parasitoid.

@@

## \*\*OP: need a section on how the food web model was fit to the data. Also should state how we deal with / use the uncertainty contained in the posterior joint distribution.

## Primary and secondary extinctions

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of three criteria: removal of (i) the most-connected species, (ii) the least-connected species and (iii) randomly chosen species. The most-connected and least-connected criteria are based on the degree (i.e. total number of links to resources and from consumers) of species. Given a primary removal if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. One no more secondary extinctions occured, then another primary extinctions was made, of the next appropriate species. This process was carried out until all species were extinct from the web.

## Calculating robustness

Robustness (R) of food web was quantified as the proportion of species subjected to primary removals that resulted in a loss (i.e. primary removals plus secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, that result in at least 50 per cent of total species loss [@dunne2002network; @dunne2004; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$), since the first primary extinction causes as cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

@@

## Show and describe the secondary extinction curves

In Fig. \ref{fig:fig_r1}, \ref{fig:fig_r2} and \ref{fig:fig_r3}, we show the secondary extinction curves of ADBM predicted food webs and observed food webs for 12 different food webs under three different extinction scenarios. We found that the cumulative secondary extinction was higher for the ADBM predicted food webs as compared to the observed food webs for nine, nine and seven food webs.

In general, irrespective of the extinction scenarios, we found that the cumulative secondary extinction was higher for the ADBM predicted food webs as compared to the observed food webs for most of the food webs **OP: this is not so clear to me. State how many cases? It seems there are many cases where secondary extinctions are more numerous in the observed food web: blue line above the red.**.

In the most connected extinction scenario, the cumulative secondary extinction curve for the observed food webs rose quickly as compared to the ADBM predicted food webs, and then reach saturation after a certain number of primary removal of species. In some of the food webs (Fig. \ref{fig:fig_r1} (f, g, h, i, j, k)), there were intersection between the cumulative secondary extinction curves of ADBM predicted food webs and that of the observed food webs. In case of the Broadstone Stream (taxonomic aggregation) food web and the Tuesday Lake food web (Fig. \ref{fig:fig_r1} (b and k)), the secondary extinction curves for the ADBM food webs were higher than the observed food webs, whereas in case of the Skipwith Pond food web (Fig. \ref{fig:fig_r1} (i)), there were no secondary extinctions for any given number of primary removal of species.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary **removals of the most connected species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_mostconnected.png")
```

Except Broadstone Stream (taxonomic aggregation), Broadstone Stream (size aggregation) and Tuesday Lake food webs (Fig. \ref{fig:fig_r2} (b, c and k)), the **mean cumulative secondary extinction curves** **OP: must be explained somewhere how/why this is a mean** for all the other food webs predicted by the ADBM were always lower than that of the observed food webs in the random extinction scenario. The shape of the cumulative secondary extinctions curves varied across the food webs.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Mean cumulative secondary extinctions of species resulting from the primary **removals of random species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_random.png")
```

Compared to the most connected and random extinction scenarios, the cumulative extinction curves in the least connected extinction scenario had very low values and were flat for most of the food webs (Fig. \ref{fig:fig_r3}) compared to the most connected extinction and random extinction scenarios. In most of the food webs, there was a lot of overlap between the extinction curves of the ADBM predicted food webs and the observed food webs.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary **removals of the least connected species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_leastconnected.png")
```

**OP: write text describing the effect sizes, e.g. how big is the difference between observed and ADBM webs.**

## Show and describe the following:

I suggest to focus on the relationship between (difference in robustness between observed and predicted connectance) and (difference in robustness between the observed and predicted food web). **OP: I miss text describing this relationship.** **OP: Allow the panels in the figure to have different y-axis scales.**

**OP: This paragraph should go in the previous section**. The ADBM predicted food webs were more robust than the observed food web with some exceptions (Fig. \ref{fig:fig_r4}). The food webs were more robust to least connected and random extinction scenarios than the primary deletion of the most connected species. The difference in the robustness values between the ADBM predicted food webs and observed food webs was higher in the most connected extinction scenario as compared to the least connected extinction and random extinction scenarios.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r4} Robustness plots for 12 food webs across different ecosystems. Here, $R_{50}$ is the proportion of species that have to be removed to achieve a total loss of at least 50\\% of total species (primary removals and secondary extinctions)."}
knitr::include_graphics("../results/plot_robustness_all_uncertainty.png")
```

AG: I will also update the robustness plot for random extinction (Fig. 4 (b)) once the simulation is completed. The one shown here is only for few simulations and as a placeholder. It takes more time to compute.

@@

The primary removal of species revealed that the ADBM predicted food webs are more robust than the observed food webs, and the shape of the robustness extinction curves varies between food webs.

The food webs are least robust to primary extinction of the most connected species scenario compared to that of least connected and random extinction scenarios. A future development would be to understand the stability of the dynamics of the ADBM predicted food webs and compare it with our study.

...

## What does overestimation of connectance imply in terms of stability?

For most of the food webs, the primary removals of species resulted in a higher secondary extinction in the ADBM predicted food webs than that of the observed food webs (Fig. \ref{fig:fig_r1}, \ref{fig:fig_r2} and \ref{fig:fig_r3}). This can be attributed to the higher connectance of the ADBM predicted food webs as compared to that of the observed food webs because a species in a food web with a high connectance has on average more number of trophic links as compared to the food webs with low connectance. **OP: should here cite figure 4.**

It would be intriguing to know if this difference in connectance has a similar influence in the dynamical stability of the food webs as well. Hence, a future prospect could be to use a dynamical model (for example: bionergetic food web model [@broseAllometricScalingEnhances2006]) to model the temporal dynamics of the ADBM predicted food webs. It would be interesting to know the temporal stability of these ADBM predicted food webs compared to the observed food webs because it has been known that food webs with increasing connectance stability diminishes [@mayWillLargeComplex1972].

...

Include here the possibility that the increased connectance would influence dynamical stability, and state if and why this may increase or decrease stability.

## Explain that the ADBM can only predicts contiguous diets [OP] and the implications of this

Using only body size as a trait, the ADBM can only predict diets that are contiguous with respect to the size of the prey. I.e. it cannot predict that a predator will consume prey of size 1 and 3, and not consume prey of size 2. Also, it is important to note that the observed diets were not contiguous when prey are ordered by their size, and this is due to some ecological differences in how predator group choose their prey [@caronAddressingEltonianShortfall]. So, the parameterisation process lead to a greater number of predicted links than observed.

This higher connectance in the ADBM predicted food webs has lead to a higher robustness of the ADBM predicted food webs. An important question to ask here is how reliable are these results. We suspect that both the model and the observed data are wrong to some extent. We expect that some of the links that do in reality occur are not present in the observed datasets, which is quite possible because of low sampling effort or rare prey-predator interactions even when there is intensive sampling. This would mean that the false positives may actually be a correctly predicted link.

>>>>

---
title: Missing links and the topological robustness of food webs
authors:
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: anubhav.gupta@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Anubhav Gupta
  thanks: Corresponding author
- affiliation: University of Zurich
  department: Department of Evolutionary Biology and Environmental Studies
  email: owen.petchey@ieu.uzh.ch
  location: 8057 Zurich, Switzerland
  name: Owen L. Petchey
geometry: margin=1in
abstract: |

  1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occuring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that predicted food webs usually have greater connectance than observed ones. Thus we expect that predicted food webs are more robust than observed ones. This expectation has never, to our knowledge, been tested, nor has the effect size been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links, though can also make false positives. We did this for 12 different food webs from a wide variety of ecosystems.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, and this can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found the food webs to be robust by 0.52 units and 0.04 units for the most connected and the random species extinction scenarios respectively while no effect in the least connected species extinction scenario.
  5)  These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, and so it would be interesting to compare the dynamical stability of observed and predicted food webs, as well as the topological stability that we have focused on.

@@

<!-- ## Background on anthropogenic changes and its impact on food web -->

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity, and can lead to food web collapse [@ullah2018]. This food web collapse is due to the cascades of secondary extinctions in a food web because of the primary loss of species, for example due to habitat destruction and climate change [@pimmHumanImpactsRates2006; @thomasComparativeLossesBritish2004; @thomasExtinctionRiskClimate2004]. Therefore, research focused on cascading secondary extinctions also known as 'community viability analysis' have been performed extensively in the past few decades to quantify how robust are food webs to species extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005]. And it has been shown that the rate of collapse of a food web is dependent on its structure and complexity [@dunne2002network; @dunneCascadingExtinctionsCommunity2009].

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. These studies have shown that the robustness of the food webs increases with food web connectance. Also, the removal of the most connected species cause considerably more secondary extinctions than the random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. Simulation studies like these which investigate the impact of primary extinctions in a food web to quantify robustness based on its topological structure provide an alternate solution to canonical experiments in natural ecosystems which are not possible or very difficult to conduct [@dunneCascadingExtinctionsCommunity2009].

Along with quantifying food web robustness based on its topological structure, studies such as @williamsEffectsNetworkDynamical2008, @broseAllometricScalingEnhances2006 and @martinezDiversityComplexityPersistence have quantified robustness based on the abundance dynamics of a food web. The topological approach of quantifying a food web robustness only requires the food web structure whereas the dynamical approach not only requires the food web structure but also the temporal dynamics of abundance of species in that food web. For example: @williamsEffectsNetworkDynamical2008 combined models of network structure with models of bioenergetic dynamics to study the role of food web topology and nonlinear dynamics on species coexistence in complex ecological networks.

<!-- ## [OP] State what is the problem... -->

A key assumption of the observed food webs is that they are very well sampled i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and then do not represent all of the feeding links that occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort as compared to others whereas some trophic links remain unobserved because of linkage constraints irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. Previous studies such as @caronAddressingEltonianShortfall and @guptaSimultaneouslyEstimatingFood2022 have shown that the predicted food webs from these models usually have greater connectance than the observed ones. Therefore, one solution to compensate for undersampling is to use a food web model such as the Allometric Diet Breadth Model (ADBM) to predict the missing links, and to then measure the robustness of the predicted food web. ADBM is a mechanistic model constructed using foraging rules based on the body sizes of prey and predator where trophic interactions satisfying those rules would be predicted by the model which are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links, and also may predict some links that could not, in fact occur.

<!-- ## What we do in this study -->

In our study, we investigate the topological robustness of the ADBM predicted food webs and compare it to that of the observed food webs, and quantify the effect of undersampling i.e overestimation of connectance on the robustness of these predicted food webs. We expect that the ADBM predicted food webs to be more robust as compared to the observed food webs, and for the greater robustness to be related to the amount by which the ADBM overestimates connectance. We do this by simulating primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

<!-- [OP] ## Provide overview of the methods -->

In the upcoming sections, we present a detailed account of the implementation of simulation of primary extinctions for three different extinction scenarios on 12 food webs predicted by the ADBM from a wide variety of ecosystems and compute the resultant secondary extinctions. We then compute a robustness metric to quantify the robustness of those predicted food webs and compare them against that of the observed food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. We chose this model because the ADBM can likely predict missing links in the predicted food webs because it consistently overestimated connectance in the predicted food webs as shown by our study in @guptaSimultaneouslyEstimatingFood2022. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables used in the model are the energy content of prey, handling times of the predator on prey, space clearance rate i.e. how fast a predator searches space, and prey densities. All of these variables are derived from the allometric scaling relationship using the body sizes of species. More details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in @petchey2008.

@@

The observed food webs that we fit the ADBM to belong to marine, freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). These food webs contain primary producers, herbivores, carnivores, parasites, and parasitoids and also contain various types of feeding interactions, including predation, herbivory, bacterivory, parasitism and pathogenic. The observed connectance of these food webs varies from 0.03 to 0.24 and the number of species varies from 29 to 239 species.

The goodness of fit of the ADBM's predictions depends on the interaction types in the food webs with predictions that are more size-structured such as aquatic and herbivory interactions being better predicted when compared to less size-structured ones such as parasitoid and terrestrial herbivory ones [@petchey2008].

@@

<!-- ## \*\*OP: need a section on how the food web model was fit to the data. Also should state how we deal with / use the uncertainty contained in the posterior joint distribution. -->

## Model parameterisation using approximate Bayesian computation

The ADBM was parameterised using approximate Bayesian computation (ABC) where a set of parameter values were sampled from the prior distributions. That set of parameter values was either accepted or rejected based on how close the predicted food web is to the observed food web using an accuracy metric -- true skill statistic (TSS). The accepted parameter values then formed a posterior distribution. Further, prediction intervals of the true skill statistic and connectance of the predicted food webs were computed. In our study, we considered food webs where the predicted connectance lay within the 95% prediction interval. A detailed explanation of the parameterisation method can be found in @guptaSimultaneouslyEstimatingFood2022.

## Extinction scenarios and robustness

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of the three criteria: removal of (i) the most-connected species, (ii) the least-connected species and (iii) randomly chosen species. The most-connected and least-connected criteria are based on the degree (i.e. the total number of links to resources and from consumers) of species. We considered these mentioned criteria because the random extinction scenario takes into account all the theoretically possible extinction sequences of species that can occur in a food web and the extinction of most-connected species and least-connected species takes into account the two opposite extreme scenarios. These extinction scenarios have been widely used in studying species extinctions and collapse of food webs and other networks [@dunne2002network; @soleComplexityFragilityEcological2001; @dunne2004; @dunneCascadingExtinctionsCommunity2009; @albertStatisticalMechanicsComplex2002].

Given a primary removal of species in a food web if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. Once no more secondary extinctions occurred, then another primary extinction was made, of the next appropriate species depending on the extinction scenario. This process was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of species subjected to primary removals that resulted in a set of extinction (i.e. primary removals plus secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, which results in at least 50% of total species loss [@dunne2002network; @dunne2004; @jonssonReliabilityR50Measure2015; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$), the first primary extinction causes a cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

## Simulating species extinctions

First, we simulated primary species loss in food webs predicted by the ADBM which had the maximum true skill statistics and compared it to primary species loss in observed food webs. Second, to take into account the uncertainty in robustness in the ADBM predicted food webs we simulated primary species loss and thereby computed robustness for all the ADBM predicted food webs corresponding to the 95% prediction interval of the predicted connectance. In the case of the random extinction scenario, we simulated 1000 random extinction sequences in a single ADBM predicted food web.

## Statistical analysis

In the random extinction scenario, we computed robustness $R_{50}$ for all 1000 independent random extinction sequences and calculated the median as a summary statistics to quantify the average robustness of a single food web. To quantify the effect of undersampling i.e. overestimation of connectance we compute the ratio of the difference in normalised robustness between the ADBM predicted food webs and observed food webs to the difference in their normalised connectance, where normalisation was performed by dividing the variables by their maximum possible values (I.e 0.5 for $R_{50}$ and 1 for connectance). However, we did not perform any statistical significance test because we work with simulated food webs and therefore the p-values of these tests are influenced by the number of model simulations [@whiteEcologistsShouldNot2014].

@@

We first present the secondary extinction curves of the ADBM predicted food webs which had the maximum true skill statistics and the observed food webs for 12 food webs under three different extinction scenarios. We then compare the robustness of all the ADBM predicted food webs within the 95% prediction interval against that of the observed food webs to take into account uncertainty in the robustness across food webs predictions. Finally we quantify the effect of overestimation of connectance on the difference in their robustness estimates.

## Secondary extinctions

In the most-connected extinction scenario, the cumulative secondary extinction curves started to rise steeply at a lower number of primary species removal in the observed food webs as compared to the ADBM predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d, e, f, g, h, j, l)). However there were higher number of cumulative secondary extinctions occurring in the ADBM predicted food webs when compared to that of the observed food webs at a high number of primary species removal (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith Pond food web, there were no secondary extinctions for any number of primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the Broadstone Stream (taxonomic aggregation) food web the same was true only for the observed food web but in the ADBM predicted food web there was a steep rise in the cumulative secondary extinctions (Fig. \ref{fig:fig_r1} (b)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of the most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_mostconnected_maxTSS.png")
```

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary extinctions in the ADBM predicted and the observed food webs for five (out of 1000) independent random extinction sequences to show qualitative differences between their secondary extinction curves. The secondary extinctions in the ADBM predicted food webs were more abrupt than that in the observed food webs.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_ra_extlines_maxTSS.png")
```

Compared to the most-connected and random extinction scenarios, there were fewer secondary extinctions in the least-connected extinction scenario and therefore the secondary extinction curves were flat for most of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs, the extinction curves of the ADBM predicted food webs overlapped with the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)}). Contrary to the most-connected scenario, there was a very high number of secondary extinctions occurring at very low number of primary species removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of the least connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_leastconnected_maxTSS.png")
```

## Robustness

The ADBM predicted food webs were more robust than the observed food webs on average in the most-connected and random extinction scenarios (Fig. \ref{fig:fig_r4} (a, b)). However, there were large variations in the robustness within the ADBM predicted food webs in the most-connected extinction scenario (Fig. \ref{fig:fig_r4} (a)). For example, the ADBM predicted Caricaie Lakes food web was more robust than the observed food web on average but had a larger variation in the robustness within the ADBM predicted food webs compared to other food webs.

The food webs were more robust to the random extinction scenario than the most-connected scenario (Fig. \ref{fig:fig_r4} (a, b)). Small Reef and Benguela Pelagic food webs had more variations in robustness within the ADBM predicted food webs as compared to the other food webs (Fig. \ref{fig:fig_r4} (b)). Skipwith Pond, Broadstone Stream (taxonomic aggregation) and Broadstone Stream (size aggregation) food webs were the most robust (Median $R_{50} = 0.5$) for both ADBM predicted and observed food webs. Although there were few less robust ADBM predicted food webs in the Broadstone Stream (size aggregation) as shown by the outliers.

In the least-connected extinction scenario, the food webs had a very high robustness (Median $R_{50} = 0.5$) for most of the food webs (Fig. \ref{fig:fig_r4} (c)), however there were some exceptions. The ADBM predicted food webs for Small Reef and Benguela Pelagic had very low median robustness. Benguela Pelagic, Broom and Capinteria food webs from the ADBM had larger variations in robustness when compared to that of the others.

```{r, echo=FALSE, fig.align='center', out.width="450px", fig.cap="\\label{fig:fig_r4} Robustness comparison between the ADBM predicted food webs and the observed food webs for 12 food webs across different ecosystems. Here, $R_{50}$ is the proportion of species that have to be removed to achieve a total loss of at least 50\\% of total species (primary removals and secondary extinctions). Box represent 25th and 75th percentile; solid diamond represent median; whisker represent outlier limits; the outlier coefficient used was 1.5. Some points are not visible due to perfect overlap in b and c. Refer to Fig. 7 in the Supplementary Information for a faceted visualisation."}
knitr::include_graphics("../results/plot_R50_ADBM_vs_obs.png")
```

In all of the food webs except Small Reef and Broadstone Stream (taxonomic aggregation), the effect size was positive on average in the most-connected extinction scenario (Fig. \ref{fig:fig_r5} (a)) i.e. overestimation of connectance had a positive effect on the robustness. In the random extinction scenario, there was a positive effect of overestimation of connectance on the robustness for Ythan, Small Reef, Mill Stream, Grasslands, Caricaie Lakes, Capinteria, Broom and Benguela Pelagic (Fig. \ref{fig:fig_r5} (b)). However, the effect size varied across the food webs. In the least-connected extinction scenario, the overestimation of connectance had little effect on average on the difference in the robustness (Fig. \ref{fig:fig_r5} (c)).

```{r, echo=FALSE, fig.align='center', out.width="450px", fig.cap="\\label{fig:fig_r5} Effect size (i.e ratio of difference in normalised robustness between ADBM predicted food webs and observed food webs to difference in their normalised connectance) shown for the 12 food webs. Box represent 25th and 75th percentile; black bold midline represent median; whisker represent outlier limits; the outlier coefficient used was 1.5."}
knitr::include_graphics("../results/plot_R50_slope.png")
```

@@

The primary removal of species revealed that the ADBM predicted food webs are more robust than the observed food webs on average, with variation within the robustness of the ADBM predicted food webs in the 12 food web ecosystems suggesting that undersampling in food webs can lead to differences in the estimates of food web robustness. The food webs are least robust to primary extinction of the most connected species scenario compared to that of least connected and random extinction scenarios on average. A future development would be to understand how undersampling i.e. overestimation of connectance influences the stability of the dynamics of the ADBM predicted food webs against that of the observed food webs, and compare it with our study.

<!-- ## What does overestimation of connectance imply in terms of stability? -->

The robustness of the ADBM predicted food webs were higher than that of the observed food webs on average (Fig. \ref{fig:fig_r4}) for all of the 12 food web ecosystems with some exceptions. This can be attributed to the higher connectance of the ADBM predicted food webs as compared to that of the observed food webs because a species in a food web with a higher connectance has on average more trophic links as compared to a food web with a lower connectance (Fig. \ref{fig:fig_r5}). Our study suggests that it is important to consider undersampling in observed food webs when computing their robustness.

The findings from our study are similar to what had been documented for other food web models [@dunneCascadingExtinctionsCommunity2009], in terms of an increase in the robustness when the connectance of a food web is increased. However, contrary to general expectations [@dunneNetworkStructureBiodiversity2002], food web robustness did not always increase with the connectance (Fig. \ref{fig:fig_r5}). The ADBM consistently predicted higher connectance than that of the observed food webs however, in some cases an increase in connectance resulted in lower robustness (Fig. \ref{fig:fig_r5}) in the predicted food web. For example: the Benguela Pelagic and Small Reef food webs were surprisingly less robust to primary extinctions on average in the least-connected extinction scenario (Fig. \ref{fig:fig_r4} (c) and \ref{fig:fig_r5} (c)). We suspect this is because the ADBM is underestimating the proportion of basal species in the predicted food webs when compared to that of the observed food webs (Fig: 6 (a) in @guptaSimultaneouslyEstimatingFood2022). As a result, these low degree basal species are the ones to be removed at an early stage in the deletion sequence thereby resulting in an earlier food web collapse in the ADBM predicted food web as compared to that of the observed food web (Fig: \ref{fig:fig_r3} (a) and (j)). This suggests that the overestimation of connectance by the ADBM resulted in a more robust food web on average but differences in the predicted food web properties such as underestimation of the proportion of basal species and overestimation of the maximum trophic level (Fig: \ref{fig:fig_a1}) counteracted that effect and led to reduced robustness. This suggests that food web properties other than connectance play an important role in determining the robustness of a food web and therefore should be also taken into account.

<!-- ## Explain how ADBM predicts diet, that the ADBM can only predicts contiguous diets [OP] and the implications of this -->
